---
title: "Formative Assessment 3"
author: "Kian Arago & Owen Santiago"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2. A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, determine the probability that:

```{r}
# Given probabilities
P_T0 <- 0.7
P_T1 <- 0.3
P_R0_given_T0 <- 0.95
P_R1_given_T0 <- 1 - P_R0_given_T0
P_R1_given_T1 <- 0.75
P_R0_given_T1 <- 1 - P_R1_given_T1
```

## (a) Probability that a 1 was received

```{r}
P_R1 <- (P_R1_given_T0 * P_T0) + (P_R1_given_T1 * P_T1)
cat("(a) Probability that a 1 was received:", P_R1, "\n")
```

## (b) Probability that a 1 was transmitted given that a 1 was received

```{r}
P_T1_given_R1 <- (P_R1_given_T1 * P_T1) / P_R1
cat("(b) Probability that a 1 was transmitted given that a 1 was received:", P_T1_given_R1, "\n")
```

## 7. There are three employees working at an IT company: Jane, Amy, and Ava, doing 10%, 30%, and 60% of the programming, respectively. 8% of Jane’s work, 5% of Amy’s work, and just 1% of Ava‘s work is in error. What is the overall percentage of error? If a program is found with an error, who is the most likely person to have written it?
```{r}
work_done <- c(Jane = 0.10, Amy = 0.30, Ava = 0.60)
work_error <- c(Jane = 0.08, Amy = 0.05, Ava = 0.01)

error_contributions <- work_done * work_error
overall_percentage_error <- sum(error_contributions) * 100
```

## Overall percentage error
```{r}
cat("Overall percentage error:", overall_percentage_error, "%\n")
```

## Percentage error per person

```{r}
percentage_errors <- error_contributions * 100
names(percentage_errors) <- names(work_done)
print(percentage_errors)
```

## Person responsible for the error

```{r}
person_error <- error_contributions / sum(error_contributions)
person_responsible <- names(which.max(person_error))
cat("Person responsible for the error:", person_responsible, "\n")
```