---
title: "Summative Assessment 1"
author: "Arago & Santiago"
date: "`r Sys.Date()`"
output: pdf_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Problem 1

```{r}
library(shiny)
```

### Define UI

```{r}
ui <- fluidPage(
  titlePanel("Defective Product Probability Calculator"),
  sidebarLayout(
    sidebarPanel(
      numericInput("x1", "Production Share of Factory 1 (x1):", 0.3, min = 0.1, max = 0.4),
      numericInput("x2", "Production Share of Factory 2 (x2):", 0.3, min = 0.1, max = 0.4),
      numericInput("x3", "Production Share of Factory 3 (x3):", 0.4, min = 0.1, max = 0.4),
      numericInput("y1", "Defective Rate of Factory 1 (y1):", 0.03, min = 0.01, max = 0.05),
      numericInput("y2", "Defective Rate of Factory 2 (y2):", 0.04, min = 0.01, max = 0.05),
      numericInput("y3", "Defective Rate of Factory 3 (y3):", 0.05, min = 0.01, max = 0.05),
      actionButton("calculate", "Calculate Probability")
    ),
    mainPanel(
      verbatimTextOutput("result")
    )
  )
)
```

### Define Server Logic

```{r}
server <- function(input, output, session) {
  observeEvent(input$calculate, {
    x1 <- input$x1
    x2 <- input$x2
    x3 <- input$x3
    y1 <- input$y1
    y2 <- input$y2
    y3 <- input$y3
    
    if (round(x1 + x2 + x3, 2) != 1) {
      output$result <- renderText("Error: x1 + x2 + x3 must equal 1.")
      return()
    }
    
    if (sum(y1, y2, y3) != 0.12) {
      output$result <- renderText("Error: y1 + y2 + y3 must equal 0.12.")
      return()
    }
    
    p_defective <- x1 * y1 + x2 * y2 + x3 * y3
    
    output$result <- renderText({
      paste("The probability of selecting a defective product is:", round(p_defective, 4))
    })
  })
}
```

### Run the Shiny App

```{r}
shinyApp(ui, server)
```

### Problem 2 

```{r}
library(shiny)
library(ggplot2)
```

### Define UI

```{r}
ui <- fluidPage(
  titlePanel("Discrete Random Variable Analysis"),
  sidebarLayout(
    sidebarPanel(
      radioButtons("var_type", "Select Variable Type:",
                   choices = c("Univariate" = "uni", "Bivariate" = "bi")),
      numericInput("num_values", "Number of Values:", 5, min = 2, max = 10),
      uiOutput("value_inputs"),
      actionButton("calculate", "Calculate")
    ),
    mainPanel(
      plotOutput("pdfPlot"),
      plotOutput("cdfPlot"),
      verbatimTextOutput("summary")
    )
  )
)
```

### Define Server Logic

```{r}
server <- function(input, output, session) {
  observeEvent(input$num_values, {
    output$value_inputs <- renderUI({
      tagList(
        lapply(1:input$num_values, function(i) {
          fluidRow(
            column(6, numericInput(paste0("val", i), paste("X", i, "Value"), value = i)),
            column(6, numericInput(paste0("prob", i), paste("P(X=", i, ")"), value = 1/input$num_values, min = 0, max = 1))
          )
        })
      )
    })
  })

  observeEvent(input$calculate, {
    values <- sapply(1:input$num_values, function(i) input[[paste0("val", i)]])
    probs <- sapply(1:input$num_values, function(i) input[[paste0("prob", i)]])

    if (sum(probs) != 1 || any(probs < 0 | probs > 1)) {
      output$summary <- renderText("Error: Probabilities must be in [0,1] and sum to 1.")
      return()
    }

    mean_value <- sum(values * probs)
    variance_value <- sum((values - mean_value)^2 * probs)

    output$summary <- renderText({
      paste("Mean:", mean_value, "\nVariance:", variance_value)
    })

    df <- data.frame(X = values, P = probs)
    output$pdfPlot <- renderPlot({
      ggplot(df, aes(x = factor(X), y = P)) +
        geom_bar(stat = "identity", fill = "blue", alpha = 0.7) +
        labs(title = "Probability Mass Function", x = "X", y = "P(X)") +
        theme_minimal()
    })

    output$cdfPlot <- renderPlot({
      df <- df[order(df$X),]
      df$CDF <- cumsum(df$P)
      ggplot(df, aes(x = factor(X), y = CDF, group = 1)) +
        geom_step(color = "red") +
        labs(title = "Cumulative Distribution Function", x = "X", y = "F(X)") +
        theme_minimal()
    })
  })
}
```

### Run the Shiny App

```{r}
shinyApp(ui, server)
```

### Problem 3

We simulate a search engine going through a list of sites for a given key phrase, stopping once the phrase is found. Each site contains the key phrase with probability `p = 0.6`.

```{r}
set.seed(123)

p <- 0.6
n_simulations <- 10000

search_counts <- rgeom(n_simulations, p) + 1

library(ggplot2)
ggplot(data.frame(X = search_counts), aes(x = X)) +
  geom_bar(aes(y = ..prop..), fill = "blue", alpha = 0.6) +
  labs(title = "Simulated Probability Density Function",
       x = "Number of Searches", y = "Probability") +
  theme_minimal()
```

### Mean and Variance Calculation

```{r}
simulated_mean <- mean(search_counts)
simulated_variance <- var(search_counts)

theoretical_mean <- 1/p

theoretical_variance <- (1-p) / (p^2)
```

```{r, echo=FALSE}
cat("Simulated Mean:", simulated_mean, "\n")
cat("Simulated Variance:", simulated_variance, "\n")
cat("Theoretical Mean:", theoretical_mean, "\n")
cat("Theoretical Variance:", theoretical_variance, "\n")
```

### Conditional Distribution Given X > 3

```{r}
conditional_data <- search_counts[search_counts > 3]

cond_mean <- mean(conditional_data)
cond_variance <- var(conditional_data)
```

```{r, echo=FALSE}
cat("Conditional Mean (X > 3):", cond_mean, "\n")
cat("Conditional Variance (X > 3):", cond_variance, "\n")
```

### Markov Memoryless Property Check

We estimate:
1. \( P(X = 4 | X > 3) \) and \( P(X = 1) \)
2. \( P(X = 5 | X > 3) \) and \( P(X = 2) \)

```{r}
# Empirical probability estimation
P_X_4_given_X_greater_3 <- mean(search_counts == 4) / mean(search_counts > 3)
P_X_1 <- mean(search_counts == 1)

P_X_5_given_X_greater_3 <- mean(search_counts == 5) / mean(search_counts > 3)
P_X_2 <- mean(search_counts == 2)
```

```{r, echo=FALSE}
cat("P(X = 4 | X > 3):", P_X_4_given_X_greater_3, "\n")
cat("P(X = 1):", P_X_1, "\n")

cat("P(X = 5 | X > 3):", P_X_5_given_X_greater_3, "\n")
cat("P(X = 2):", P_X_2, "\n")
```

This confirms the Markov memoryless property, as the conditional probabilities align with the unconditional geometric probabilities.
